Modal Abstractions for Virtualizing Memory Addresses

### 2025-07-28 programlogic.tex
more cleanup and compression, now halfway through 4.2

\add{
  A program logic for reasoning about code which may work with (and possibly update) multiple address spaces requires dealing with
  several key challenges. It must ensure that reasoning about memory accesses only depends on assertions that hold
  in the active address space at the time of access. It must allow invariants for code or data structures to refer
  to other address spaces. These constraints mean it must also support reasoning about when the active address space
  changes, as this affects which memory assumptions are usable or not for memory access and thus which data structures are
  immediately directly accessible.
}
\add{
  Many approaches which could handle these problems in principle, such as tagging pointers with the relevant address space.
  But such approaches introduce other complexities.
  Most code, even in an operating system kernel, only works with a single address space (the current one).
  Explicitly plumbing address space information through a simple linked list specification, simply because some other part of the program
  may manipulate other address spaces, adds significant specification burden.
  Specifications which need to talk about a particular invariant holding in an arbitrary other address space,
  or with knowing a certain assertion conceptually holds in multiple specific address spaces, would need to quantify
  over functions from address space identity to assertions rather than just assertions.
  Using a modal approach resolves all of these technical and usability\todo{need better word, I'm always wary of using this word in PL}
  challenges cleanly and uniformly.
  Specifications that are not \emph{about} address space manipulation need not mention an address spaces at all:
  one can use one's favorite separation logic data structure specifications without restructuring,
  and every assertion can be stated for either the current or specific other address spaces.
  In short, modalities make it possible for specifications to mention address spaces only when they deal with address spaces,
  and not when they don't.\todo{wary of this due to the rtv in the double}
}

@@

% The idea that the truth of an assertion is relative to an address space has far-reaching consequences.
To support making assertions depend on a choice of address space, we work entirely in a pointwise lifting of \iris's base BI logic,
essentially working with separation logic assertions indexed by a choice of page table root as a $\mathcal{W}_{64}$, which we call $\textsf{vProp }\Sigma$:\footnote{
  \iris experts may notice our \lstinline|-b>| resembles another pointwise lifting already in  \iris~\cite{dang2019rustbelt,dang2022compass}. 
  This similarity is real, but the existing lifting does not appear to work with indexed \coq types like our \lstinline|word n| as a domain.
}
\lstinline[language=Coq]|Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$|.
This is the (\rocq) type of assertions in our logic.
Most constructs in \iris's base logic are defined with respect to any BI-algebra (of \coq type \lstinline|bi|), so \add{they} automatically
carry over to our derived logic.
\add{Since \textsf{vProp}s are ``just'' \textsc{Iris} predicates on a choice of page table root,
it may not be obvious what the gain is over simply using \textsc{Iris} directly.
The gain is not expressive power.
Every modal logic can be translated straightforwardly to a logic without modalities
(nearly always a small variation on the \emph{standard translation} from propositional modal logic
to first-order logic~\cite{blackburn2006modal}\footnote{\add{Itself similar to verification condition generation
for weakest precondition calculi and Hoare logics.}}) that can directly address the logic's models ---
but these translations, in addition to making specifications far more verbose,
generally do not preserve the appealing reasoning principles of modal logics.
This becomes \replace{most}{more} apparent when discussing our proof rules.
}

However, we must still build up from existing \iris primitives to provide new primitives that depend on the address space --- primarily the notion
of virtual points-to.
To define and use virtual points-to assertions, we require two basic assertions that ignore
the current address space:

\paragraph{Register points-to} 
The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ containing the
value of the register $\rv$.\todo{Ismail, you changed ``the register value $\rv$'' to ``value of the register $\rv$'' and I changed it back. What you wrote treats $\rv$ as the name of the register, not its value.}
The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping and grants update permission \replace{on}{to} it;
otherwise, any value $0 < \qfrac <1$ represents partial ownership, granting read-only permission on the mapping.

@@

 %I put a newline in between the following two Definitions as the second Definition seems not proper without the newline
\begin{figure*}
  \begin{lstlisting}[language=Coq]
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
  $\exists_{\textsf{l4e l3e l2e l1e}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\vaddr$ l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
 Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ := $\lambda$ cr3val.
  $\ulcorner$ entry_present l4e $\land$ entry_present l3e $\land$ entry_present l2e $\land$ entry_present l1e$\urcorner$ $\ast$
  (l4M52 maddr cr3val) $\sim$ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$
  (l3M52 maddr l4e) $\sim$ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
  (l2M52 maddr l3e) $\sim$ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e $\ast$ $\ulcorner$ addr_L1(va,l1e) = paddr $\urcorner$.
\end{lstlisting}
\todo[inline,color=cyan]{Ismail, we can drop the maddr from the l4M52 etc, and the l4e etc. from the l3off etc., right?}
\vspace{-1em}
\caption{A Strong Virtual Points-to Relation
}
  \label{fig:strongvirtualpointsto}
\vspace{-1em}
\end{figure*}

@@

\todo[inline]{This is a section where it's worth explaining the figure in a little more detail, especially as we reuse L4L1PointsTo
in our own definitions later. I know we discussed at some point, but don't the \texttt{l$n$M52} accessors ignore the \texttt{maddr},
and don't the \texttt{l$n$off} accessors ignore the entry arguments? If so we can drop those (simplifying the figure), and annotate 
(e.g., with comments like \lstinline[mathescape]|(*$\textcircled{E}$*)|) each \texttt{l$n$off} in Figure \ref{fig:strongvirtualpointsto}
with a tag matching a new tag on the corresponding offset chunk of the translated virtual addres in Figure \ref{fig:pagetables}\ldots
}
\todo[inline]{Ismail Addressed: current explanation in here is fortified by the changed text in the background section smoothly}

\lstinline|L4_L1_PointsTo| works by
chaining together the entries for each level, using the sequence of table offsets from the address being translated to index
each table level, and using the physical page address embedded in each entry.\footnote{
  The fractions \lstinline|q1| through \lstinline|q4| represent the fractional ownership of each entry based on how many
  word-aligned addresses might need to share the entry ---  $(\frac{1}{512})^n$ for each level $n$.
}
For example, the first-level address translation to get the L4 entry (\lstinline|l4e|) 
  uses the masks \textsf{l4M52} with the current \lstinline|cr3| to get the physical address of the start of the L4 table
  and \textsf{l4off} with the virtual address being translated to compute the correct offset within that table.
  Then \del{at} that physical location \replace{is}{holds} the appropriate entry in the L4 table \textsf{l4e}.
  Subsequent levels of the page table walk work similarly.
The statement of these assertions is simplified by the use of our split physical points-to assertions, since
each level of tables is page-sized. \footnote{We do not address superpages and hugepages in this paper.}
This helper definition is also more explicit in \textsf{vProp} which binds a value to \lstinline|cr3| and uses it to start the translation process.
\looseness=-1

@@

This solution is in fact very close to that of \citet{kolanski08vstte}, who define a separation logic from scratch in \textsc{Isabelle/HOL},
where the semantics of all assertions are functions from pairs of heaps and page table root values to booleans.\footnote{
  This was a typical explicit construction at the time; their work significantly predates \iris.
}
Our solution \replace{removes}{in the next subsection improves on theirs by removing} some restrictions in this definition by further abstracting the handling of address translation.

@@

  We separate the physical page-table walk from the virtual points-to relation, replacing it with a ghost state that merely guarantees that the address translation would succeed.
  \iris includes a \emph{ghost map} construction, which we use to track mappings from virtual addresses to the physical addresses they translate to as a piece of ghost state.
  The map includes, for each key in the map (i.e.,
  each virtual address), a token $\ghostmaptoken{\gamma}{k}{v}$ whose ownership is required to update that key-value pair in the ghost map named $\gamma$. The existence of such a token implies that the actual map $\theta$ tracked by a corresponding $\mathsf{GhostMap}(\gamma,\theta)$
  resource indeed maps $k$ to $v$. These properties are captured by key \iris rules in Figure \ref{fig:ghostmaps}.\footnote{\iris ghost maps lack established notation in the literature,
  but the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}
  There are other rules, but these two are most important for explaining ghost maps.
  \textsc{GhostMapUpdate} says that ownership of the actual ghost map with ghost name $\gamma$ and map contents $\theta$,
  and a full (non-fractional) token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
  changing the map and replacing the token to represent the new value.
  \textsc{GhostMapLookup} allows using the same information to simply conclude that the mapping indicated by the token is true.

@@

  % In Iris this is realized by using an authoritative resource algebra: there is a single \emph{authoritative} global copy of the (ghost)
  % map caching virtual-to-physical address translations, and for each entry a read only \emph{partial} ownership of that key-value pair.
  % The resource algebra itself is instantiated as:
  % \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
For clarity, we refer to the specific ghost map summarizing virtual-to-physical translations by \\
\centerline{$\mathcal{A}\textsf{bsPTableWalk}(\delta,\theta) \stackrel{\triangle}{=} \mathsf{GhostMap}(\delta,\theta)$}\\
(omitting $\delta$ for brevity when only one is in scope)
and keep this in a per-address-space invariant described shortly.
We then replace the physical traversal $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}'s virtual points-to definition
with ownership of the token \ghostmaptoken{\delta}{\vaddr}{\paddr}, %ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
yielding Figure \ref{fig:virtualpointstosharing}'s definition.
This new definition guarantees that the ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$),
and thus that the per-address-space invariant will contain the physical resources that guarantee \add{that} the hardware resolves the translation.

  \begin{figure*}
  \footnotesize
\centerline{$
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast \\
 \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \textsf{ where } 
   \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
    %\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}
    \ptableabswalk{\delta,\theta}
\end{array}
$}
\vspace{-1em}
\caption{Per-address-space invariant with a fixed global map of address space names $m$}
  \label{fig:peraspaceinvariant}
\vspace{-1em}
  \end{figure*}

We place the authorative ownership of the ghost translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
$\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}). 
$\mathcal{I}$\textsf{ASpace} alone allows changes to the page tables
that preserve overall virtual-to-physical translations.
When combined with the
fragment stored in the virtual points-to (Figure \ref{fig:virtualpointstosharing}),
changes in the virtual-to-physical translations become possible without the need to track all virtual points-to assertions.

\todo[inline]{This next paragraph below explains details, but should first explain the big picture: the current address space
is identified by a paddr, so to state the invariant for the address space named by that paddr we need to look up
what invariants should hold, then assert that those invariants do hold.}
In Figure \ref{fig:peraspaceinvariant},
$\theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
for the \emph{current} address space, corresponding to the  currently installed page tables (the ones
indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially quantified $\delta$
is the ghost name for this map (again, for the current address space).
Because our logic is the first to address the existence of multiple address spaces,
it must ensure that $\theta$ actually corresponds to the intended ghost map for the current address space.
This requires looking up the ghost name (i.e., $\delta$) for the current address space,
which is handled by $m$ --- another logical map (tied to ghost state shortly)
from various page table roots to ghost names for the per-address-space ghost maps discussed above.
$\mathcal{I}\textsf{ASpace}(\theta,m)$ then stores the physical resources for the current address space mappings,
ensuring that the current address space is valid (has an entry in $m$ whose corresponding
authoritative map matches $\theta$), and asserting ownership of the appropriate fractional ownership
of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).

\todo[inline]{What's not clear in this next paragraph is the relationship between ASpaceLookup in Figure 5 and the lookup part of Figure 3.
I think Figure 5's $m$ and Figure 3's $\delta{s}$ are the same, aren't they? If so, we should say so and use the same name.
If not we should explain why they're different.
}
Virtual points-to assertions must also ensure that they are tied to a valid address space, and because they
use ghost map tokens to ensure the virtual-to-physical translation exists in a specific ghost map, they must also
name that map. In Figure \ref{fig:virtualpointstosharing}, this is done with a ghost map token for \emph{another} ghost map,
one shared across the kernel, holding $m$. Our virtual points-to assertion existentially quantifies the address-space-specific
ghost map $\delta$ mapping virtual addresses to physical addresses, and ensures it exists with a token for the distinguished
global (unique) ghost map named by $\delta{s}$.
% the ghost map of virtual-to-physical translations for a particular address space (the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}).
% % for each address space must be locateable from a representation of the specific address space:
% % This means we also require a way to locate the map for the current address space from 
% We support lookup by the current root page table pointer stored in \lstinline|cr3|.
\looseness=-1

>>>>

\add{
  A program logic for reasoning about code which may work with (and possibly update) multiple address spaces requires dealing with
  several key challenges. It must ensure that reasoning about memory accesses only depends on assertions that hold
  in the active address space at the time of access. It must allow invariants for code or data structures to refer
  to other address spaces. These constraints mean it must also support reasoning about when the active address space
  changes, as this affects which memory assumptions are usable or not for memory access and thus which data structures are
  immediately directly accessible.
}
\add{
  Many approaches which could handle these problems in principle, such as tagging pointers with the relevant address space.
  But such approaches introduce other complexities.
  Most code, even in an operating system kernel, only works with a single address space (the current one).
  Explicitly plumbing address space information through a simple linked list specification, simply because some other part of the program
  may manipulate other address spaces, adds significant specification burden.
  Specifications which need to talk about a particular invariant holding in a specific other address space (or multiple other address spaces) would need to quantify
  over \emph{functions} from address space identity to assertions rather than just assertions.
  Using a modal approach resolves all of these challenges cleanly and uniformly.
  Specifications that are not \emph{about} address space manipulation need not mention an address spaces at all:
  one can use standard separation logic data structure specifications without restructuring,
  and every assertion can be stated for either the current or specific other address spaces.
  In short, modalities make it possible for specifications to mention address spaces when necessary,
  without requiring mention when only one address space is at hand.
}

@@

% The idea that the truth of an assertion is relative to an address space has far-reaching consequences.
To support making assertions depend on a choice of address space, we work entirely in a pointwise lifting of \iris's base BI logic,
essentially working with separation logic assertions indexed by a choice of page table root as a $\mathcal{W}_{64}$, which we call $\textsf{vProp }\Sigma$:\footnote{
  \iris experts may notice our \lstinline|-b>| resembles another pointwise lifting already in  \iris~\cite{dang2019rustbelt,dang2022compass}. 
  This similarity is real, but the existing lifting does not appear to work with indexed \coq types like our \lstinline|word n| as a domain.
}
\lstinline[language=Coq]|Definition vProp  $\Sigma$ : bi := word 64 -b> iPropI  $\Sigma$|.
This is the (\rocq) type of assertions in our logic.
Most constructs in \iris's base logic are defined with respect to any BI-algebra (of \coq type \lstinline|bi|), so \add{they} automatically
carry over to our derived logic.
% \add{Since \textsf{vProp}s are ``just'' \textsc{Iris} predicates on a choice of page table root,\todo{this section may be subsumed under first addition to this section, don't need both}
% it may not be obvious what the gain is over simply using \textsc{Iris} directly.
% The gain is not expressive power.
% Every modal logic can be translated straightforwardly to a logic without modalities
% (nearly always a small variation on the \emph{standard translation} from propositional modal logic
% to first-order logic~\cite{blackburn2006modal}\footnote{\add{Itself similar to verification condition generation
% for weakest precondition calculi and Hoare logics.}}) that can directly address the logic's models ---
% but these translations, in addition to making specifications far more verbose,
% generally do not preserve the appealing reasoning principles of modal logics.
% This becomes \replace{most}{more} apparent when discussing our proof rules.
% }
However, we must still build up from existing \iris primitives to provide new primitives that depend on the address space --- primarily the notion
of virtual points-to.
To define and use virtual points-to assertions, we require two basic assertions that ignore
the current address space:

\paragraph{Register points-to} 
The assertion $\textsf{r}\;\mapsto_{\textsf{r}}\{q\}\;\textsf{rv}$ ensures the ownership of the register $\rg$ containing the
value of the register $\rv$.
The fraction $\qfrac$ with value 1 asserts the unique ownership of the register mapping and grants update permission \replace{on}{to} it;
otherwise, any value $0 < \qfrac <1$ represents partial ownership, granting read-only permission on the mapping.

@@

 %I put a newline in between the following two Definitions as the second Definition seems not proper without the newline
\begin{figure*}
  \begin{lstlisting}[language=Coq,escapeinside={(*}{*)}]
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
  $\exists_{\textsf{l4e l3e l2e l1e}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo($\vaddr$ l4e l3e l2e l1e paddr) $\ast$ paddr $\mapsto_{p}\{\mathsf{q}\} \vpage$.
 Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e paddr :word 64) : vProp $\Sigma$ := $\lambda$ cr3val.
  $\ulcorner$ entry_present l4e $\land$ entry_present l3e $\land$ entry_present l2e $\land$ entry_present l1e$\urcorner$ $\ast$
  (l4M52 maddr cr3val) $\sim$ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$(*\label{line:l4pointsto}*)
  (l3M52 maddr l4e) $\sim$ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$(*\label{line:l3pointsto}*) 
  (l2M52 maddr l3e) $\sim$ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$(*\label{line:l2pointsto}*)
  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e $\ast$ $\ulcorner$ addr_L1(va,l1e) = paddr $\urcorner$.(*\label{line:l1pointsto}*)
\end{lstlisting}
\todo[inline,color=cyan]{Ismail, we can drop the maddr from the l4M52 etc, and the l4e etc. from the l3off etc., right?}
\vspace{-1em}
\caption{A Strong Virtual Points-to Relation
}
  \label{fig:strongvirtualpointsto}
\vspace{-1em}
\end{figure*}

@@

\lstinline|L4_L1_PointsTo| works by
chaining together the entries for each level, using the sequence of table offsets from the address being translated to index
each table level, and using the physical page address embedded in each entry.\footnote{
  The fractions \lstinline|q1| through \lstinline|q4| represent the fractional ownership of each entry based on how many
  word-aligned addresses might need to share the entry ---  $(\frac{1}{512})^n$ for each level $n$.
}
For example, the first-level address translation to get the L4 entry (\lstinline|l4e|) 
  uses the masks \textsf{l4M52} with the current \lstinline|cr3| to get the physical address of the start of the L4 table
  and \textsf{l4off} with the virtual address being translated to compute the correct \add{byte} offset within that table \add{just as in the first translation
  step of Figure \ref{fig:pagetables}}.\footnote{\add{Note the offsets mentioned in Figure \ref{fig:pagetables} are 9-bit indexes into the 512 entries; the byte offset is that times 8.}}
  \replace{
    Then {at} that physical location {is} the appropriate entry in the L4 table \textsf{l4e}.
  }{
    Thus Line \ref{line:l4pointsto} asserts that the physical address built from the table base and offset points to the L4 entry \textsf{l4e}.
  }
  Subsequent levels of the page table walk \add{assertion (Lines \ref{line:l3pointsto}--\ref{line:l1pointsto})} work similarly.
The statement of these assertions is simplified by the use of our split physical points-to assertions, since
each level of tables is page-sized. \footnote{We do not address superpages and hugepages in this paper.}
This helper definition is also more explicit in \textsf{vProp} which binds a value to \lstinline|cr3| and uses it to start the translation process.
\looseness=-1

@@

This solution is in fact very close to that of \citet{kolanski08vstte}, who define a separation logic from scratch in \textsc{Isabelle/HOL},
where the semantics of all assertions are functions from pairs of heaps and page table root values to booleans.\footnote{
  This was a typical explicit construction at the time; their work significantly predates \iris.
}
Our solution \replace{removes}{in the next subsection improves on theirs, removing} some restrictions in this definition by further abstracting the handling of address translation.
\looseness=-1

@@

  We separate the physical page-table walk from the virtual points-to relation, replacing it with a ghost state that merely guarantees that the address translation would succeed.
  \iris includes a \emph{ghost map} construction, which we use to track mappings from virtual addresses to the physical addresses they translate to as a piece of ghost state.
  The map includes, for each key in the map (i.e.,
  each virtual address), a token $\ghostmaptoken{\gamma}{k}{v}$ whose ownership is required to update that key-value pair in the ghost map named $\gamma$. The existence of such a token implies that the actual map $\theta$ tracked by a corresponding $\mathsf{GhostMap}(\gamma,\theta)$
  resource indeed maps $k$ to $v$. These properties are captured by key \iris rules in Figure \ref{fig:ghostmaps}.\footnote{\iris ghost maps lack established notation in the literature,
  but the syntax we use captures the details of \texttt{iris.base\_logic.lib.ghost\_map}.}
  There are other rules, but these two are most important for explaining ghost maps.
  \textsc{GhostMapUpdate} says that ownership of the actual ghost map with ghost name $\gamma$ and map contents $\theta$,
  and a \del{full (non-fractional) }token witnessing that $\theta$ maps \textsf{pa} to \textsf{va} permits an update to the ghost map's state,
  changing the map and replacing the token to represent the new value.
  \textsc{GhostMapLookup} allows using the same information to simply conclude that the mapping indicated by the token is true.

@@

  % In Iris this is realized by using an authoritative resource algebra: there is a single \emph{authoritative} global copy of the (ghost)
  % map caching virtual-to-physical address translations, and for each entry a read only \emph{partial} ownership of that key-value pair.
  % The resource algebra itself is instantiated as:
  % \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
For clarity, we refer to the specific ghost map summarizing virtual-to-physical translations by 
\mbox{$\mathcal{A}\textsf{bsPTableWalk}(\delta,\theta) \stackrel{\triangle}{=} \mathsf{GhostMap}(\delta,\theta)$}
(omitting $\delta$ for brevity when only one is in scope)
and keep this in a per-address-space invariant described shortly.
We then replace the physical traversal $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}'s virtual points-to definition
with ownership of the token \ghostmaptoken{\delta}{\vaddr}{\paddr}, %ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
yielding Figure \ref{fig:virtualpointstosharing}'s definition.
This new definition guarantees that the ghost map \add{$\theta$} \replace{contains an mapping from}{maps} the virtual address ($\vaddr$) to a physical address ($\paddr$),
and thus that the per-address-space invariant \add{described next} will contain the physical resources that guarantee \add{that} the hardware resolves the translation.
\looseness=-1

  \begin{figure*}
  \footnotesize
\vspace{-1em}
\centerline{$
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{=} \textsf{ASpace\_Lookup}(\ptablestore,m) \ast \\
 \qquad\qquad\qquad \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \textsf{ where } 
   \textsf{ASpace\_Lookup} (\ptablestore,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
    %\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}
    \ptableabswalk{\delta,\theta}
\end{array}
$}
\vspace{-1em}
\caption{Per-address-space invariant with a fixed global map of address space names $m$}
  \label{fig:peraspaceinvariant}
\vspace{-1em}
  \end{figure*}

We place the authorative ownership of the ghost \add{map} translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
$\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}). 
$\mathcal{I}$\textsf{ASpace} alone allows changes to the page tables
that preserve overall virtual-to-physical translations.
When combined with the
fragment stored in the virtual points-to (Figure \ref{fig:virtualpointstosharing}),
changes in the virtual-to-physical translations become possible without the need to track all virtual points-to assertions.

\todo[inline,color=cyan]{Explain $\delta{s}$ (Iris ghost name) vs $m$ (logical contents of ghost map) in next 2 paragraphs}
% \todo[inline]{This next paragraph below explains details, but should first explain the big picture: the current address space
% is identified by a paddr, so to state the invariant for the address space named by that paddr we need to look up
% what invariants should hold, then assert that those invariants do hold.}
\add{One remaining difficulty is ensuring that different address spaces can have independent ghost maps ---
which we resolve with an additional global ghost map from address-space identifiers (page table roots) to
the \textsc{Iris} ghost name for that address space's ghost map.}
In Figure \ref{fig:peraspaceinvariant},
$\theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
for the \emph{current} address space, corresponding to the  currently installed page tables (the ones
indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially quantified $\delta$
is the ghost name for this map (again, for the current address space).
Because our logic is the first to address the existence of multiple address spaces,
it must ensure that $\theta$ actually corresponds to the intended ghost map for the current address space.
This requires looking up the ghost name (i.e., $\delta$) for the current address space,
which is handled by $m$ --- another logical map (tied to ghost state shortly)
from various page table roots to ghost names for the per-address-space ghost maps discussed above.
$\mathcal{I}\textsf{ASpace}(\theta,m)$ then stores the physical resources for the current address space mappings,
ensuring that the current address space is valid (has an entry in $m$ whose corresponding
authoritative map matches $\theta$), and asserting ownership of the appropriate fractional ownership
of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).

Virtual points-to assertions must also ensure that they are tied to a valid address space, and because they
use ghost map tokens to ensure the virtual-to-physical translation exists in a specific ghost map, they must also
name that map. In Figure \ref{fig:virtualpointstosharing}, this is done with a ghost map token for \emph{another} ghost map,
one shared across the kernel, holding $m$. Our virtual points-to assertion existentially quantifies the address-space-specific
ghost map $\delta$ mapping virtual addresses to physical addresses, and ensures it exists with a token for the distinguished
global (unique) ghost map named by $\delta{s}$.
% the ghost map of virtual-to-physical translations for a particular address space (the ghost name $\delta$ in Figure \ref{fig:virtualpointstosharing}).
% % for each address space must be locateable from a representation of the specific address space:
% % This means we also require a way to locate the map for the current address space from 
% We support lookup by the current root page table pointer stored in \lstinline|cr3|.
\looseness=-1
